---
title: "Reproducable Reasearch Coursework 1"
author: "William Moore"
date: "February 5, 2016"
output: html_document
---

# Read the data
The data set of qualified self movement was initially downloaded from [here](https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip) on 2016-02-05.  
First I read the data

```{r}
rawData <- read.csv("../Data/raw_data.csv")
```

# Total Number of Steps Each Day
The histogram below shows a histogram of the total number of steps taken each day

```{r}
stepsOnDay <- subset(rawData, !is.na(rawData$steps), select=c("date", "steps"))
stepsOnDay$date <- as.Date(stepsOnDay$date)

totalStepsOnDay <- aggregate(stepsOnDay[c("steps")], by=list(stepsOnDay$date), FUN=sum, 1)

hist(totalStepsOnDay$steps, main = "Total Steps", xlab = "Days with Count Total Steps", ylab = "Number of Days", breaks = 20)

meanStepsOnDay <- mean(totalStepsOnDay$steps)
medianStepsOnDay <- median(totalStepsOnDay$steps)
```

Mean steps = `r meanStepsOnDay`  
Median steps = `r medianStepsOnDay` 


# Average Daily Pattern
```{r}
stepsInteavl <- subset(rawData, !is.na(rawData$steps), select=c("interval", "steps"))

averageStepsInteval <- aggregate(stepsInteavl[c("steps")], by=list(stepsInteavl$interval), FUN=mean, 1)

plot(averageStepsInteval$Group.1, averageStepsInteval$steps, type="l", main = "Average Daily Pattern", xlab = "Time", ylab = "Steps")

maxInteval <- which.max(averageStepsInteval$steps)
maxIntevalTime <- averageStepsInteval$Group.1[maxInteval]
maxIntevalValue <- averageStepsInteval$steps[maxInteval]
```

The inteval with the most steps was `r maxIntevalTime`. There were average steps of `r maxIntevalValue`.  

# Missing Values
Here we simulate the missing values in the original data set using the average value from that time inteval on all other days.  

```{r}
totalMissingValues <- length(rawData$steps) - length(stepsInteavl$steps)

dataWithSimulatedValues <- rawData

for (i in 1:length(dataWithSimulatedValues$steps))
{
    if(is.na(dataWithSimulatedValues$steps[i])){
        index <- Position(function(y){y == dataWithSimulatedValues$interval[i]}, averageStepsInteval$Group.1)
        dataWithSimulatedValues$steps[i] <- averageStepsInteval$steps[index]
    }
}
```

The total missing values from the original data set was `r totalMissingValues`  
We now plot with the simulated values

```{r}
stepsOnDay <- subset(dataWithSimulatedValues, select=c("date", "steps"))
stepsOnDay$date <- as.Date(stepsOnDay$date)

totalStepsOnDay <- aggregate(stepsOnDay[c("steps")], by=list(stepsOnDay$date), FUN=sum, 1)

hist(totalStepsOnDay$steps, main = "Total Steps", xlab = "Days with Count Total Steps", ylab = "Number of Days", breaks = 20)

meanStepsOnDay <- mean(totalStepsOnDay$steps)
medianStepsOnDay <- median(totalStepsOnDay$steps)
```

Mean steps = `r meanStepsOnDay`  
Median steps = `r medianStepsOnDay` 

# Differance Between Weekdays and Weekends
```{r}
dataWithSimulatedValues$IsWeekend <- weekdays.Date(as.Date(dataWithSimulatedValues$date)) == "Saturday" | weekdays.Date(as.Date(dataWithSimulatedValues$date)) == "Sunday"

weekendData <- subset(dataWithSimulatedValues, dataWithSimulatedValues$IsWeekend, select=c("interval", "steps"))
weekdayData <- subset(dataWithSimulatedValues, !dataWithSimulatedValues$IsWeekend, select=c("interval", "steps"))

averageStepsIntevalWeekend <- aggregate(weekendData[c("steps")], by=list(weekendData$interval), FUN=mean, 1)
averageStepsIntevalWeekday <- aggregate(weekdayData[c("steps")], by=list(weekdayData$interval), FUN=mean, 1)

par(mfrow=c(2,1))
plot(averageStepsIntevalWeekend$Group.1, averageStepsIntevalWeekend$steps, type="l", main = "Weekend Daily Pattern", xlab = "Time", ylab = "Steps")
plot(averageStepsIntevalWeekday$Group.1, averageStepsIntevalWeekday$steps, type="l", main = "Weekday Daily Pattern", xlab = "Time", ylab = "Steps")
```
